---
title: "P8106 Data Science II Final Project"
author: "Stephanie Zhen, Xinran Luo, Ran An"
date: "5/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message=FALSE}
library(tidyverse)
library(readr)
library(Amelia)
library(DataExplorer)
library(caret)
library(pROC)
library(ROCR)
library(klaR)
library(AppliedPredictiveModeling)
library(naivebayes)
```

age: age in years

sex: (1 = male; 0 = female)

cp: chest pain type (typical angina, atypical angina, non-angina, or asymptomatic angina)

trestbps: resting blood pressure (in mm Hg on admission to the hospital)

chol: serum cholestoral in mg/dl

fbs: Fasting blood sugar (< 120 mg/dl or > 120 mg/dl) (1 = true; 0 = false)

restecg: resting electrocardiographic results (normal, ST-T wave abnormality, or left ventricular hypertrophy)

thalach: Max. heart rate achieved during thalium stress test

exang: Exercise induced angina (1 = yes; 0 = no)

oldpeak: ST depression induced by exercise relative to rest

slope: Slope of peak exercise ST segment (0 = upsloping, 1 = flat, or 2 = downsloping)

ca: number of major vessels (0-3) colored by flourosopy 4 = NA

thal: Thalium stress test result 3 = normal; 6 = fixed defect; 7 = reversable defect 0 = NA

target: Heart disease status 1 or 0 (0 = heart disease 1 = asymptomatic)

### Load the data

```{r message=FALSE}
heart = read_csv("./data/heart.csv")
head(heart)

summary(heart)

missmap(heart)
```

###Data Cleaning:

```{r}
heart_nomiss = heart %>%
  filter(thal != 0 & ca != 4) %>% 
  mutate(
    sex = case_when(
     sex == 1 ~ "male",
    sex == 0 ~ "female"
           ),
    fbs = case_when(
      fbs == 1 ~ "true",
      fbs == 0 ~ "false"
            ),
    exang = case_when(
      exang == 1 ~ "yes",
      exang == 0 ~ "no"
            ),
    cp = case_when(
      cp == 3 ~ "typical_angina",
      cp == 1 ~ "atypical_angina",
      cp == 2 ~ "non_anginal",
      cp == 0 ~ "asymptomatic_angina"
          ),
    restecg = case_when(
      restecg == 2 ~ "wave_abnormality",
      restecg == 1 ~ "normal",
      restecg == 0 ~ "hypertrophy"
              ),
    target = case_when(
      target == 1 ~ "asymptomatic",
      target == 0 ~ "heart_disease"
              ),
    slope = case_when(
      slope == 2 ~ "upsloping",
      slope == 1 ~ "flat",
      slope == 0 ~ "downsloping"
    ),
    thal = case_when(
      thal == 3 ~ "reversable_defect",
      thal == 2 ~ "normal",
      thal == 1 ~ "fixed_defect"
    ),
    sex = as.factor(sex),
    fbs = as.factor(fbs),
    exang = as.factor(exang),
    cp = as.factor(cp),
    slope = as.factor(slope),
    ca = as.factor(ca),
    thal = as.factor(thal)
  )

```


```{r}
plot_density(heart_nomiss)
```

```{r}
plot_bar(heart_nomiss)
```


###Partition the data for training

```{r}
set.seed(1)
train.indices<-createDataPartition(y=heart_nomiss$target,p=0.7,list=FALSE)

training<-heart_nomiss[train.indices,]
testing<-heart_nomiss[-train.indices,]
```


###Naive Bayes (NB)
```{r}
set.seed(1)

heart.nb = naive_bayes(as.factor(target) ~ ., data=training)
nb.pred <- predict(heart.nb, newdata = testing, type = "prob")[,2]
roc.nb <- roc(testing$target, nb.pred)
#ROC curve
plot(roc.nb)
roc.nb$auc
```

